---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Read in the `activity.csv` file:

```{r}
library(lubridate)
activity = read.csv('activity.csv', na.strings = "NA")
activity$date <- ymd(activity$date)
head(activity)
```


## What is mean total number of steps taken per day?

```{r}
library(dplyr)
daily.activity <- group_by(activity, date) %>% summarize(total = sum(steps))
```

1. Histogram of the total number of steps taken each day 
(choosing a `binwidth=1000` based on experience with my Fitbit Surge...)

```{r}
library(ggplot2)
qplot(daily.activity$total, binwidth=1000)
```

We see that hitting 10'000 steps a day (and above) was achieved quite often. Yay! I wish *my* Fitbit stats looked like that...


2. *mean* and *median* total number of steps taken each day

Let's add some more statistical punch to the observation above... 

- The mean total number of steps taken each day was 
```{r} 
mean(daily.activity$total, na.rm=TRUE)
``` 
- and the median was 
```{r}
median(daily.activity$total, na.rm=TRUE)
```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

We need to re-group the data, this time per interval:

```{r}
interval.activity <- group_by(activity, interval) %>% summarize(avg.steps = mean(steps, na.rm = TRUE))
```

Then we can plot the time series:

```{r}
qplot(interval, avg.steps, data=interval.activity, geom='line')
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
filter(interval.activity, avg.steps==max(avg.steps)) %>% select(interval)
```

Looks like at interval 835. Hm... I wonder what time that is... say we added `835 / 5` minutes to midnight...


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
